Course Project Practical Machine Learning
=========================================

```{r, echo=FALSE}
xmsg <- sprintf("date and time: %s and version used: %s", Sys.time(), R.Version()$version.string)
```

Synopsis
--------

This prediction project was done on `r xmsg` by Lucho Farje. Using data which has been generated by devices such as Nike, FuelBand, FitBit and Jawbone Up about personal activity and self movements which were attached in their body of different persons. The goal using this dataset is to find patterns using machine learning to build a prediction model. The data to be used is coming from accelerometers on the belt, forearm, arm and dumbell of 6 participants which were asked to perform barbell lifts correctly and incorrectly in five different ways.

Setting Global Environment
--------------------------
This prediction analysis needs some packages in order to be processed. The **caret** package is important for this prediction model as well as another support packages which makes the entire analysis process much easier to perform.
An important hint in this process is to highlight that in this step it's being initialized the seed variable and creating a clusters of cores for this computer. 

```{r cache=FALSE,echo=TRUE,results='hide'}
library(caret)
library(parallel)
library(doParallel)
library(randomForest)
library(ggplot2)
library(xtable)
#using parallel clusters for the cores of this machine/server
cluster.cores<-makeCluster(detectCores()-1)
registerDoParallel(cluster.cores)
#Initialize seed
day.of.begin.project<-"2015-01-16"
seed.number<-as.numeric(as.Date(day.of.begin.project))
set.seed(seed.number)
```

Download and Load Datasets
--------------------------

In this step the data will be downloaded and loaded into variables in order to be filtered and pre-processed for applying the prediction model.

```{r, cache=TRUE, echo=TRUE}
if (!file.exists("./data/pml-testing.csv"))
  { download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "./data/pml-testing.csv") }

if (!file.exists("./data/pml-training.csv"))
  { download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "./data/pml-training.csv") }

data.training<-read.csv("./data/pml-training.csv", na.strings=c("NA",""))
data.testing<-read.csv("./data/pml-testing.csv", na.strings=c("NA",""))
```


Data Pre-Processing
-------------------

When you click the **Knit HTML** button a web page will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```

Train The Prediction Model
--------------------------





Applying Prediction Mode On Data Test
-------------------------------------



